INTEGRAL TRANSPORT CORRECTION TO DIFFUSION CALCULATIONS\\USING pCMFD SCHEME

### 2019-11-05 docs/PHYSOR2020/pCMFD-RM-PHYSOR2020.tex
updated manuscript and bib file. still no results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[letterpaper]{physor2020}
%
%  various packages that you may wish to activate for usage 
\usepackage{tabls}
\usepackage{cites}
\usepackage{epsf}
\usepackage{appendix}
\usepackage{ragged2e}
\usepackage[top=1in, bottom=1.in, left=1.in, right=1.in]{geometry}
\usepackage{enumitem}
\setlist[itemize]{leftmargin=*}
\usepackage{caption}
\captionsetup{width=1.0\textwidth,font={bf,normalsize},skip=0.3cm,within=none,justification=centering}
\usepackage{bm}
\usepackage{tikz}
\usepackage{xfrac}
\usepackage{units}

@@

%
% Define title...
%
\title{INTEGRAL TRANSPORT CORRECTION TO DIFFUSION CALCULATIONS\\USING pCMFD SCHEME}
%
% ...and authors
%
\author{%
  % FIRST AUTHORS 
  %
  \textbf{Roy Gross$^1$, Daniele Tomatis$^2$, and Erez Gilad$^3$}\footnote{} \\
  $^1$Name of Institution 1  \\
  Corresponding Address \\ 
\\
  $^2$Name of Institution 2  \\ 
    Corresponding Address \\ 
\\
  $^3$Name of Institution 3  \\
     Corresponding Address \\
     \\
  \url{roygross@post.bgu.ac.il}, \url{daniele.tomatis@cea.fr}, \url{gilade@bgu.ac.il}
}
%
% Insert authors' names and short version of title in lines below
%
\newcommand{\authorHead}      % Author's names here use et al. if more than 3
           {Roy Gross, Daniele Tomatis, Erez Gilad}  
\newcommand{\shortTitle}      % Short title here (Shorten to fit all into a single line)
           {INTEGRAL TRANSPORT CORRECTION TO ONE-DIMENSIONAL DIFFUSION CALCULATIONS}  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%   BEGIN DOCUMENT
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\maketitle
\justify

@@

The main hypothesis of this research, called the Ronen Method (RM)~\cite{Ronen-2004,Tomatis-2011}, is based on iterative calculations of the multigroup diffusion coefficients, driven by the accurate relations between the neutron current density and the neutron flux as derived from the integral transport equation. The iterative scheme calculates new flux distribution using a diffusion solver but with a (spatially) modified diffusion coefficient according to the current calculated using the integral transport equation. However, the direct resolution of the integral relation between the diffusion coefficient and the transport current would imply the inversion of large matrices, with poor control of their conditioning. Nonetheless, the integral current can be enforced in the Coarse Mesh Finite Difference (CMFD) discretized form of the diffusion equation, which avoids the numerical issues resulting from small flux gradients. This is the option adopted in this study.

% ------------------------------------------
\subsection{Coarse Mesh Finite Difference (CMFD)} 
\label{sec:CMFD-intro}

@@

% ------------------------------------------
\section{The Ronen method} 
\label{sec:RM}

In 2004, Ronen~\cite{Ronen-2004} suggested to derive corrected diffusion coefficients using Fick's law and more accurate estimations of the neutron currents by means of integral transport operators, whereas the neutron flux is resolved by diffusion theory. Denoting the integral transport and diffusion currents by $\bm{J}\rE$ and $\bm{J}_D\rE$, respectively, the Ronen idea is based on adapting the diffusion coefficient such that
\begin{equation}\label{eq:Fick}
\bm{J}^D\rE = -D\rE\bnabla\phi\rE = \bm{J}\rE \,\, . 
\end{equation}

@@

The motivation for this method was to overcome the inherent limitation of Fick's law requiring smooth flux gradients and thus weakly absorbing medium. Nevertheless, isotropic scattering remained as a basic postulate. For example, in a one-dimensional homogeneous slab with void boundary conditions, Eq.~\eqref{eq:RM-it} takes the following form~\cite{Ronen-2004} 
\begin{equation}\label{eq:RM-it-1D-slab}
D^{(k+1)}(x,E) = -\frac{\frac{1}{2}\int_0^a dx' E_2[\sigma(E)\lvert
	x-x'\rvert]\text{sign}(x-x')q^{(k)}(x',E)}
{\partial \phi^{(k)}(x,E)/\partial x} \,\, .
\end{equation}

@@

% ------------------------------------------
\section{Numerical implementation of the Ronen method} 
\label{sec:RM-num}

The CMFD neutron balance equation for a $\keff$-eigenvalue problem on a one-dimensional mesh, shown in Fig.~\ref{fig:mesh1D}, is
\begin{equation}\label{eq:balance}
J_{i+\hzi,g} - J_{i-\hzi,g} + \sigma_g \phi_{i+\hzi,g} = q_{i+\hzi,g} \,\, , 
\end{equation}	
where $J$ denotes the accurate net current at the cell surface, $\sigma$ denotes the total cross section, the subscript $g$ denotes the energy group, and $q$ is the isotropic neutron source defined as
\begin{equation}\label{eq:qsrc}
q_{i+\hzi,g} = \sum\limits_{g'=1}^G\sigma_{s,0,g\leftarrow g'}\phi_{i+\hzi,g'} +
\frac{\chi_g}{2\keff}\sum\limits_{g'=1}^G\nu\sigma_{f,i+\hzi,g'}\phi_{i+\hzi,g'} \,\, ,
\end{equation}	
where $\sigma_0$ is the isotropic scattering cross section, $\chi$ is the fission spectrum, $\keff$ is the eigenvalue, $\nu$ is the average number of neutrons emitted per thermal fission, and $\sigma_f$ is the fission cross section.

@@

	\draw ( 8.5 cm,2pt) -- ( 8.5 cm,-2pt) node[anchor=north] {$I-3/2$};
	\node[anchor=south] at (9.5cm,1pt) {$I-1$};
	\draw ( 10.5 cm,2pt) -- ( 10.5 cm,-2pt) node[anchor=north] {$I-\hzi$};
	\end{tikzpicture}
	\caption{Notation of the one-dimensional mesh.}
	\label{fig:mesh1D}
\end{figure}

% ------------------------------------------
\subsection{CMFD implementation} 
\label{sec:RM-CMFD}

% ------------------------------------------
\subsection{pCMFD implementation} 
\label{sec:RM-pCMFD}

The partial currents in any point $x$ in a slab are defined according to
\begin{equation}
\label{eq:partial-current}
\jpm (x) \cong \frac{\phi(x)}{4} \pm \frac{\jD (x)}{2} \,\, ,
\end{equation}
where $\jD$ denotes the diffusion current.
%where
%\begin{equation}
%\label{eq:diff-current}
%\bm{J} (x) = \jp(x) - \jm(x) \,\, .
%\end{equation}
The partial currents correction factors are defined on cell surface according to
\begin{equation}
\label{eq:PCCF}
\jpm _{i+\hzi} = \frac{1}{4}\phi_{i+\hzi}
\pm \frac{1}{2} \jD _{i+\hzi} \pm \frac{1}{2} \delta \jpm _{i+\hzi} \,\, ,
\end{equation}
where
\begin{equation}
\label{eq:jdiff}
\jD _{i+\hzi} = - 2 D_{i+\hzi} \frac{\phi_{i+1} - \phi_i}{\Delta_{i+1}+\Delta_i} 
\,\,, %\nonumber \\
D_{i+\hzi} = \frac{\Delta_i D_i + \Delta_{i+1} D_{i+1}}{\Delta_i + \Delta_{i+1}} 
\,\,, %\nonumber \\
\phi_{i+\hzi} = \frac{\Delta_i \phi_i + \Delta_{i+1} \phi_{i+1}}{\Delta_i + \Delta_{i+1}} \, .
\end{equation}
Substituting~\cite{Jarrett-2016,Zhu-2016}
\begin{equation}
\label{eq:CCF-sub}
\frac{1}{2} \delta \jp _{i+\hzi} \equiv -\delta D^+ _{i+\hzi} \phi_i \,, 
\quad
\frac{1}{2} \delta \jm _{i+\hzi} \equiv -\delta D^- _{i+\hzi} \phi_{i+1}\,\, ,
\end{equation}
yields
\begin{align}
\label{eq:par4}
\jp _{i+\hzi} &= \frac{1}{4}\phi_{i+\hzi}
+ \frac{1}{2}\jD _{i+\hzi} - \delta D^+ _{i+\hzi} \phi_i \\
\jm _{i+\hzi} &= \frac{1}{4}\phi_{i+\hzi}
- \frac{1}{2}\jD _{i+\hzi} + \delta D^- _{i+\hzi} \phi_{i+1} \,\, .
\end{align}
%Recall that $\jpm$ are calculated using the integral expression. 
%
Solving for the partial current correction factors gives
\begin{equation}
\label{eq:ccfsss1}
\delta D^+ _{i+\hzi} = \frac{\frac{1}{4}\phi_{i+\hzi}
	+ \frac{1}{2}\jD _{i+\hzi} - \jp _{i+\hzi}}{\phi_i} \,, \quad
\delta D^- _{i+\hzi} = \frac{-\frac{1}{4}\phi_{i+\hzi}
	+ \frac{1}{2}\jD _{i+\hzi} + \jm _{i+\hzi}}{\phi_{i+1}} \,\, .
\end{equation}

@@

% ------------------------------------------
\section{Results} 
\label{sec:res}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[letterpaper]{physor2020}
%
%  various packages that you may wish to activate for usage 
\usepackage{tabls}
\usepackage{cite}
\usepackage{epsf}
\usepackage{appendix}
\usepackage{ragged2e}
\usepackage[top=1in, bottom=1.in, left=1.in, right=1.in]{geometry}
\usepackage{enumitem}
\setlist[itemize]{leftmargin=*}
\usepackage{caption}
\captionsetup{width=1.0\textwidth,font={bf,normalsize},skip=0.3cm,within=none,justification=centering}
\usepackage{bm}
\usepackage{tikz}
\usepackage{xfrac}
\usepackage{units}

@@

%
% Define title...
%
\title{INTEGRAL TRANSPORT CORRECTION TO DIFFUSION CALCULATIONS\\USING pCMFD SCHEME}
%
% ...and authors
%
\author{%
  % FIRST AUTHORS 
  %
  \textbf{Roy Gross$^1$, Daniele Tomatis$^2$, and Erez Gilad$^{1,}$}\footnote{Corresponding author} \\
  $^1$The Unit of Nuclear Engineering \\
  Ben-Gurion University of the Negev, 8410501 Beer-Sheva, Israel\\ 
\\
  $^2$CEA, DEN, Service d’\'etudes des r\'eacteurs et
  de math\'ematiques appliqu\'ees (SERMA) \\
  Universit\'e Paris-Saclay, F-91191, Gif-sur-Yvette, France \\ 
\\
  \url{roygros@post.bgu.ac.il}, \url{daniele.tomatis@cea.fr}, \url{gilade@bgu.ac.il}
}
%
% Insert authors' names and short version of title in lines below
%
\newcommand{\authorHead}      % Author's names here use et al. if more than 3
           {R. Gross, D. Tomatis, E. Gilad}  
\newcommand{\shortTitle}      % Short title here (Shorten to fit all into a single line)
           {The Ronen Method by pCMFD}  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%   BEGIN DOCUMENT
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\maketitle
\justify

@@

%The main hypothesis of this research, called the Ronen Method (RM)~\cite{Ronen-2004,Tomatis-2011}, is based on iterative calculations of the multigroup diffusion coefficients, driven by the accurate relations between the neutron current density and the neutron flux as derived from the integral transport equation. The iterative scheme calculates new flux distribution using a diffusion solver but with a (spatially) modified diffusion coefficient according to the current calculated using the integral transport equation. However, the direct resolution of the integral relation between the diffusion coefficient and the transport current would imply the inversion of large matrices, with poor control of their conditioning. Nonetheless, the integral current can be enforced in the Coarse Mesh Finite Difference (CMFD) discretized form of the diffusion equation, which avoids the numerical issues resulting from small flux gradients. This is the option adopted in this study.

% ------------------------------------------
%\subsection{Coarse Mesh Finite Difference (CMFD)} 
%\label{sec:CMFD-intro}

@@

% ------------------------------------------
\section{THE RONEN METHOD} 
\label{sec:RM}

In 2004, Ronen~\cite{Ronen-2004} suggested to derive corrected diffusion coefficients using Fick's law and more accurate estimations of the neutron currents by means of integral transport operators, whereas the neutron flux is resolved by diffusion theory. Denoting the integral transport and diffusion currents by $\bm{J}\rE$ and $\bm{J}^D\rE$, respectively, the Ronen idea is based on adapting the diffusion coefficient such that
\begin{equation}\label{eq:Fick}
\bm{J}^D\rE = -D\rE\bnabla\phi\rE = \bm{J}\rE \,\, . 
\end{equation}

@@

The motivation for this method was to overcome the inherent limitation of Fick's law requiring smooth flux gradients and thus weakly absorbing medium. Nevertheless, isotropic scattering remained as a basic postulate. For example, in a one-dimensional homogeneous slab with void boundary conditions, Eq.~\eqref{eq:RM-it} takes the following form~\cite{Ronen-2004} 
\begin{equation}\label{eq:RM-it-1D-slab}
D^{(k+1)}(x,E) = -\frac{\frac{1}{2}\int_0^a dx' E_2[\sigma(E)\lvert
	x-x'\rvert]\text{sign}(x-x')q^{(k)}(x',E)}
{\partial \phi^{(k)}(x,E)/\partial x} \,\, ,
\end{equation}
where the source term $q^{(k)}(x',E)$ is calculated using the multigroup diffusion model with available diffusion coefficients, and the updated diffusion coefficients $D^{(k+1)}(x,E)$ are evaluated using the integral expression for the transport current.

@@

% ------------------------------------------
\section{NUMERICAL IMPLEMENTATION} 
\label{sec:RM-num}

The CMFD neutron balance equation for a $\keff$-eigenvalue problem on a one-dimensional mesh, shown in Fig.~\ref{fig:mesh1D}, is
\begin{equation}\label{eq:balance}
J_{i+\hzi,g} - J_{i-\hzi,g} + \sigma_{i,g} \phi_{i,g} = q_{i,g} \,\, , 
\end{equation}	
where $J$ denotes the accurate net current at the cell surface, $\sigma$ denotes the total cross section, the subscript $g$ denotes the energy group, and $q$ is the isotropic neutron source defined as
\begin{equation}\label{eq:qsrc}
q_{i,g} = \sum\limits_{g'=1}^G\sigma_{s,0,i,g\leftarrow g'}\phi_{i,g'} +
\frac{\chi_g}{2\keff}\sum\limits_{g'=1}^G\nu\sigma_{f,i,g'}\phi_{i,g'} \,\, ,
\end{equation}	
where $\sigma_0$ is the isotropic scattering cross section, $\chi$ is the fission spectrum, $\keff$ is the eigenvalue, $\nu$ is the average number of neutrons emitted per thermal fission, and $\sigma_f$ is the fission cross section.

@@

	\draw ( 8.5 cm,2pt) -- ( 8.5 cm,-2pt) node[anchor=north] {$I-\sfrac{3}{2}$};
	\node[anchor=south] at (9.5cm,1pt) {$I-1$};
	\draw ( 10.5 cm,2pt) -- ( 10.5 cm,-2pt) node[anchor=north] {$I-\hzi$};
	\end{tikzpicture}
	\caption{Notation of the one-dimensional mesh.}
	\label{fig:mesh1D}
\end{figure}

The accurate net current and partial current in the case of isotropic scattering are calculated using the following expressions derived from the integral neutron transport equation~\cite{Tomatis-2011,Tomatis-2019,Gross-2020}
\begin{align}\label{eq:par-curr-calc}
\jp_{i+\hzi,g} &=  \frac{1}{2}
E_3[\mcL_g(x_{-\hzi},x_{i+\hzi})] \phi_{x_{-\hzi},g} \nonumber \\
&+
\frac{1}{2}\sum\limits_{j=0}^i \frac{q_{j,g}}{\sigma_{j,g}}
\left\{ 
E_3\left[
\mcL_g(x_{j+\hzi},x_{i+\hzi}) 
\right]
- E_3\left[
\mcL_g(x_{j-\hzi},x_{i+\hzi})
\right] 
\right\} 
\nonumber \\
\jm_{i+\hzi,g} &=  \frac{1}{2}
E_3[\mcL_g(x_{i+\hzi},x_{I-\hzi})] \phi_{x_{I-\hzi},g} \nonumber \\
&+
\frac{1}{2}\sum\limits_{j=i+1}^{I-1} \frac{q_{j,g}}{\sigma_{j,g}}
\left\{ 
E_3\left[
\mcL_g(x_{i+\hzi},x_{j-\hzi}) 
\right]
- E_3\left[ 
\mcL_g(x_{i+\hzi},x_{j+\hzi})
\right] 
\right\} \,\, ,
\end{align}
where $\mcL_g(x',x)\equiv \int_{x'}^{x} \sigma_g(x'')dx''$ is the optical length and the net current is $J _{i+\hzi,g} = \jp _{i+\hzi,g} - \jm _{i+\hzi,g}$.

% ------------------------------------------
\subsection{CMFD IMPLEMENTATION} 
\label{sec:RM-CMFD}

In standard CMFD implementation, correction terms, e.g., $\delta J_{i+\hzi,g}$, are added to the diffusion current $\jD_{i+\hzi,g}$, such that it is driven towards the accurate net current $J_{i+\hzi,g}$. The accurate net current is calculated by the integral expression given in~Eq.~\ref{eq:par-curr-calc} and using the scalar flux for evaluating the emission source $q_{i,g}$. The diffusion current is evaluated according to 
\begin{equation}\label{eq:JD}
\jD_{i+\hzi,g} \cong -D_{i+\hzi,g} 
\frac{\phi_{i+1,g} - \phi_{i,g}}
{(\Delta_{i+1} + \Delta_i)/2} \,\, ,
\end{equation}
where integer and rational subscripts indicate node-averaged and interface quantities, respectively, and 
\begin{equation}\label{eq:Ds}
D_{i+\hzi,g} = \frac{\Delta_i D_{i,g} + \Delta_{i+1} D_{i+1,g}}
{\Delta_i + \Delta_{i+1}} \, \, .
\end{equation}
%
Once the accurate net current and the diffusion current are calculated using the most recent flux values, the correction terms on cell interfaces can be calculated. The discretized form of the current $\delta J_{i+\hzi,g}$ must involve the neighboring flux as well, but its representation is changed into a drift-advection term to avoid of possible undefined division by zeros in case of flat flux~\cite{Smith-1983,Tomatis-2011}
\begin{equation}
\delta J_{i+\hzi,g} = -\delta D_{i+\hzi,g} 
\frac{\phi_{i+1,g} + \phi_{i,g}}
{(\Delta_{i+1} + \Delta_i)/2} 
= J_{i+\hzi,g} - \jD_{i+\hzi,g}
\,\, .
\end{equation}

% ------------------------------------------
\subsection{pCMFD IMPLEMENTATION} 
\label{sec:RM-pCMFD}

The partial currents in any point $x$ in a slab are defined according to
\begin{equation}
\label{eq:partial-current}
\jpm (x) \cong \frac{\phi(x)}{4} \pm \frac{\jD (x)}{2} \,\, ,
\end{equation}
where $\jD$ denotes the diffusion current.
%where
%\begin{equation}
%\label{eq:diff-current}
%\bm{J} (x) = \jp(x) - \jm(x) \,\, .
%\end{equation}
The partial currents correction factors are defined on cell surface according to
\begin{equation}
\label{eq:PCCF}
\jpm _{i+\hzi} = \frac{1}{4}\phi_{i+\hzi}
\pm \frac{1}{2} \jD _{i+\hzi} \pm \frac{1}{2} \delta \jpm _{i+\hzi} \,\, ,
\end{equation}
where
\begin{equation}
\label{eq:jdiff}
\phi_{i+\hzi} = \frac{\Delta_i \phi_i + \Delta_{i+1} \phi_{i+1}}{\Delta_i + \Delta_{i+1}} \, .
\end{equation}
Substituting~\cite{Jarrett-2016,Zhu-2016}
\begin{equation}
\label{eq:CCF-sub}
\frac{1}{2} \delta \jp _{i+\hzi} \equiv -\delta D^+ _{i+\hzi} \phi_i \,, 
\quad
\frac{1}{2} \delta \jm _{i+\hzi} \equiv -\delta D^- _{i+\hzi} \phi_{i+1}\,\, ,
\end{equation}
yields
\begin{align}
\label{eq:par4}
\jp _{i+\hzi} &= \frac{1}{4}\phi_{i+\hzi}
+ \frac{1}{2}\jD _{i+\hzi} - \delta D^+ _{i+\hzi} \phi_i \\
\jm _{i+\hzi} &= \frac{1}{4}\phi_{i+\hzi}
- \frac{1}{2}\jD _{i+\hzi} + \delta D^- _{i+\hzi} \phi_{i+1} \,\, .
\end{align}
%Recall that $\jpm$ are calculated using the integral expression. 
%
Solving for the partial current correction factors gives
\begin{equation}
\label{eq:ccfsss1}
\delta D^+ _{i+\hzi} = \frac{\frac{1}{4}\phi_{i+\hzi}
	+ \frac{1}{2}\jD _{i+\hzi} - \jp _{i+\hzi}}{\phi_i} \,, \quad
\delta D^- _{i+\hzi} = \frac{-\frac{1}{4}\phi_{i+\hzi}
	+ \frac{1}{2}\jD _{i+\hzi} + \jm _{i+\hzi}}{\phi_{i+1}} \,\, .
\end{equation}

@@

% ------------------------------------------
\section{RESULTS} 
\label{sec:res}
